---
title: "linear models 1"
author: "Laura Robles-Torres"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type)
```

#Fitting models 

Fitting initial model of price (y - outcome) on rating "stars" (x1) and borough (x2) with this code:

The lm function begins with the formula specification – outcome on the left of the ~ and predictors separated by + on the right. As we’ll see shortly, interactions between variables can be specified using *. You can also specify an intercept-only model (outcome ~ 1), a model with no intercept (outcome ~ 0 + ...), and a model using all available predictors (outcome ~ .).
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```

Let's look at the result. Does not give you a dataframe but does provide a object of class lm. Used the bronx as a reference group given categorical variable borough. 

x0 = -70.41
x1 = 31.99
x2 = 40.50
x3 = 90.25
x4 = 13.21

```{r, include=FALSE}
fit
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
```
'summary' gives you residual information, coefficient estimates, std errors, and runs t-test for each coefficient, as well as R^2, adjusted R^2, the F-test 
if you add '$coef' it will give you just the info re: to the coefficients.
coef(fit) will give you just the coefficients.

But...none of these are tidy. We use the broom package for that.The broom package has functions for obtaining a quick summary of the model and for cleaning up the coefficient table. 

```{r}
fit |> 
  broom::glance() 

fit |>
  broom::tidy()

```
Both of these functions *produce data frames*, which makes it straightforward to include the results in subsequent steps.

```{r}
fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(term = str_replace(term, "^borough", "Borough: ")) |> 
  knitr::kable(digits = 3)
```

#Diagnostics 

```{r}

```

